Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2019-06-13 papers/lime_diagnostics/visual-diagnostics-of-a-model-explainer.Rnw
worked on intro

\hh{Generally: avoid 'can be'. Replace by 'is'.  }

@@

\kg{Added in Heike's suggestions and made a few changes of my own to this paragraph:} In the field of statistics, there are two main uses for models: interpretation and prediction. Machine learning models are often used for the latter purpose. While these models have been proven to provide accurate predictions in a wide range of problems, this predictive ability comes at the cost of interpretability. These models are typically complex and lack a functional form, which makes it not possible to directly interpret the models and has earned them the reputation of being black boxes. The ability to interpret a model helps to diagnose the model and understand the predictions, and in some areas of application, the need to be able to explain the results from a model is critical.

The forensics sciences and the heath care industry are two areas that have benefitted from machine learning but have an obvious need for explainable predictions. For example, \citet{aoas} discuss the use of a random forest model in the forensics sciences to determine whether two bullets were fired from the same gun. \citet{yu:2018} describe how healthcare has advanced through the use of machine learning in ways such as using neural networks to automate medical image diagnoses and implementing Bayesian networks to predict clinical outcomes. In \hh{both} of these cases, the decisions made based on the results of the machine learning models would greatly impact the lives of human beings. If it is not possible to explain the output from a model, it becomes questionable whether to trust the results from the model to make decisions.

@@

The European Union has taken the desire to provide explanations from machine learning models a step further. In May 2018, the General Data Protection Regulation (GDPR) went into effect (https://eugdpr.org/the-regulation/gdpr-faqs/). The regulation includes a policy that gives the right for subjects affected by decisions made via automated algorithms to understand the logic behind the decisions being made. As Goodman and Flaxman (2016) point out, this regulation will have a great effect on the way that machine learning algorithms are used to make decisions. At minimum, it will require that an explanation of how the input variables relate to the predictions.

\kg{Moved the last sentence from the first paragraph here:} As a result of the desire to be able to explain the predictions, an area of research has emerged on machine learning explain the output from machine learning algorithms. One approach that is being taken to remedy the problem of uninterpretable models is to develop a model explainer that provides insight into the performance of the complex model. A model explainer is a method that is separate from the model but uses the model and output from the model to provide understanding into mechanism of how the model produces predictions. Recently, a handful of model explainers have been developed (reference). LIME (local interpretable model explanations) is one such model explainer (reference).

@@

The explanation that is produced from the explainer is also a prediction or an element of a model. Since it is not directly extracted from the complex model, it is important to ask the question, ``How reliable is the explanation?". Currently, there are no recommendations for how to assess the explanations from LIME (need to check to make sure this is a true statement). \kg{I'd like to expand on this paragraph since this is really the focus of the paper. Then the last paragraph in the intro can outline the paper.}

>>>>

To check on when editing writing: 
\begin{itemize}
\item \hh{Generally: avoid 'can be'. Replace by 'is'.  }
\item \kg{words to go back and make sure I am not using too often: ability, produce}
\end{itemize}

@@

\kg{I added in Heike's suggestions to the first paragraph and made some changes of my own. I divided the first paragraph into two parts and added a discusssion on reasons why interpretability is beneficial. Note that I also switched the two model uses from ``interpretation and prediction" to ``inference and prediction". Since a key part of the paper is about being able to interpret predictive models, I thought it might be good to distinguish between interpretation and inference. Heike, what do you think?}\\
\\
In the field of statistics, there are two main uses for models: \kg{inference} and prediction. Machine learning models are often used for the latter purpose. While these models have been proven to perform well in a wide range of problems, their predictive accuracy comes at the cost of interpretability. They are often complex algorithms that are good at identifying patterns in the data but lack a functional form, so it not possible to directly interpret them. As a result, they produce predictions that lack an explanation, which has earned them the reputation of being black box models.

The ability to interpret a model serves multiple purposes. When a model is first fit, interpretations help to diagnosing the model. If it is possible to know what variables are influencing a prediction, it can be determined if the predictions are based on reasonable variables. If not, appropriate adjustments to the model can be made. After a model has been diagnosed and found to be reasonable, the explanations for the predictions can be used to understand the underlying mechanism that produced the data and provide an argument for why the predictions should be trusted. In some areas of application, this ability to be able to explain the results from a model is critical.

The forensics sciences and the heath care industry are two areas that have benefitted from machine learning but have an obvious need for explainable predictions. For example, \citet{aoas} discuss the use of a random forest model in the forensics sciences to determine whether two bullets were fired from the same gun. \citet{yu:2018} describe how healthcare has advanced through the use of machine learning in ways such as using neural networks to automate medical image diagnoses and implementing Bayesian networks to predict clinical outcomes. In both of these \kg{fields}, the decisions made based on the results of the machine learning models would greatly impact \kg{people's lives}. If it is not possible to explain the output from a model, it becomes questionable whether to \kg{rely} on the predictions to make important decisions.

@@

The European Union has taken the desire to provide explanations from machine learning models a step further. In May 2018, the General Data Protection Regulation (GDPR) went into effect \footnote{https://eugdpr.org/the-regulation/gdpr-faqs/}. The regulation includes a policy that gives the right for \kg{individuals} affected by decisions made via automated algorithms to understand the logic behind the decisions being made. As \citet{goodman:2016} point out, this regulation will have a great effect on the way that machine learning algorithms are used to make decisions. \kg{While they point out that the wording of the policy leaves room for interpretation}, they believe that, at minimum, it will require that an explanation of how the input variables relate to the predictions.

\kg{As a result of the push (wording?) to produce explainable predictions, an area of research has emerged focusing on developing ways to explain output from machine learning algorithms.} One approach that is being taken to remedy the problem of uninterpretable models is to develop model explainers that provide insight into the performance of the complex model. A model explainer is a method that is separate from the model but uses the model and output from the model to provide understanding into mechanism of how the model produces predictions. Recently, a handful of model explainers have been developed (reference). LIME (local interpretable model explanations) is one such model explainer (reference).

@@

The explanation that is produced from the explainer is also a prediction or an element of a model. Since it is not directly extracted from the complex model, it is important to ask the question, ``How reliable is the explanation?". Currently, there are no recommendations for how to assess the explanations from LIME (need to check to make sure this is a true statement). \kg{LIME provides easily understandable explanations, but the current implementations of LIME are based on linear models, which are not assessed.} \kg{I'd like to expand on this paragraph since this is really the focus of the paper. Then the last paragraph in the intro can outline the paper.}
